---
title: "SDS 315 - Homework 9"
author: "Amanda Roberts: acr4437"
date: "April 16, 2024"
output:
  pdf_document: default
  html_document: default
---

Github: https://github.com/amandarob8/SDS-315/tree/main/Homework%209

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(mosaic)
library(MatchIt)
library(moderndive)
library(kableExtra)
```


# Problem 1: Get Out The Vote
```{r include=FALSE}
vote = read.csv("turnout.csv")
```

## Part A
```{r include=FALSE}
#prop of those receiving/not receiving a GOTV call who voted in 1998
prop(voted1998 ~ GOTV_call, data=vote, success=1)
#yes call(1): 0.6477733
#no call(0): 0.4442449

#confidence interval
prop.test(voted1998 ~ GOTV_call, data=vote, success=1)
#95 interval: -0.2659167, -0.1411399
#yes call more likely to vote (by 20ish %)
```

The proportion of those receiving a "get out the vote"(GOTV) call who voted in 1998 is about 0.648, and the proportion of those who did not receive a GOTV call who also voted in 1998 is about 0.444. A large sample 95% confidence interval for the difference in these two proportions is about -0.266 to -0.141, indicating those who received a call were more likely to vote in 1998 by a difference somewhere in that interval, 95% of the time.


## Part B
```{r include=FALSE}
#compare confounders with treatment(GOTV_call)

#yes call: 0.7125506    no call: 0.5308070
prop(voted1996 ~ GOTV_call, data=vote)

#yes call: 58.30769     no call: 49.42534
mean(AGE ~ GOTV_call, data=vote)

#yes call: 0.8016194    no call: 0.7447552
prop(MAJORPTY ~ GOTV_call, data=vote)

#clear differences in call rate...could contribute to a difference in response (voted1998)
```

The variables voted1996, AGE, and MAJORPTY could be confounders if they would make someone more likely to receive a GOTV call and have voted in 1998, preventing the difference observed above from representing the true difference. The variable voted1996, whether the person voted in 1996, presents the proportions 0.713 for those who received a call and 0.531 for those who did not. The variable AGE, the person's age, presents the means 58.3 for those who received a call and 49.4 for those who did not. The variable MAJORPTY, whether the person is registered as one of the two major U.S. political parties, presents the proportions 0.802 for those who received a call and 0.745 for those who did not. All three variables have a larger mean or proportion for those who received a call, making them confounders in this situation.


## Part C
```{r include=FALSE}
#confounder: factors that make someone more likely to receive a call and have voted in 1998
vote_matched = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data=vote, ratio=5) %>% 
  match.data

#yes call: 0.7125506      no call: 0.7125506
diffprop(voted1996 ~ GOTV_call, data=vote_matched)

# yes call: 58.30769   no call: 58.26640
diffmean(AGE ~ GOTV_call, data=vote_matched)

#yes call: 0.8016194      no call: 0.8072874
diffprop(MAJORPTY ~ GOTV_call, data=vote_matched)
#all balanced
```

```{r include=FALSE}
#prop of those receiving/not receiving a GOTV call who voted in 1998
prop(voted1998 ~ GOTV_call, data=vote_matched, success=1)
#yes call(1): 0.6477733
#no call(0): 0.5740891

#confidence interval
prop.test(voted1998 ~ GOTV_call, data=vote_matched, success=1)
#95 interval: -0.141756693, -0.005611728
#yes call slightly more likely to vote
```

After using the matching technique with GOTV_call as the treatment variable and voted1996, AGE, and MAJORPTY as the matching variables, the variables are not longer confounders since matched data set is balanced. The difference in proportions between call and no call for the variable voted1996 is 0.000. The difference in means between call and no call for the variable AGE is 0.041. The difference in proportions between call and no call for the variable MAJORPTY is -0.006. Since the differences are insignificant, the new matched data set is balanced with respect to the three variables.


With the matched data set, the recalculated proportion of those receiving a "get out the vote"(GOTV) call who voted in 1998 is about 0.648, and the proportion of those who did not receive a GOTV call who also voted in 1998 is about 0.574. A large sample 95% confidence interval for the difference in these two proportions is about -0.142 to -0.006, indicating those who received a call are now only slightly more likely to vote in 1998 by a difference somewhere in that interval, 95% of the time. Based on these calculations, the GOTV call did produce a slight increase in people voting in 1998, after controlling for known confounders.


\newpage
# Problem 2: Manufacturing Flaws
```{r include=FALSE}
solder = read.csv("solder.csv")
```

## Part A

Graphed below are the relationships between the number of skips on a circuit board, manufacturing flaws, and either the opening size of the solder gun or the thickness of the alloy used for soldering. In the first plot, it is clear that smaller the opening size, the larger the number of skips, on average. The second plot depicts that a thinner alloy creates, on average, more skips.

```{=tex}
\vspace{12pt}
```

```{r include=FALSE}
#opening vs skips
solder$Opening[solder$Opening == "S"] = "Small"
solder$Opening[solder$Opening == "M"] = "Medium"
solder$Opening[solder$Opening == "L"] = "Large"

solder$Opening = factor(solder$Opening, levels=c("Small", "Medium", "Large"))

opening_plot = ggplot(solder) +
  geom_jitter(aes(x=Opening, y=skips), width=0.1, col="royalblue3") +
  theme_bw() +
  labs(title="Size of Opening vs Number of Skips",
       x="Size of Opening",
       y="Number of Skips") +
  stat_summary(aes(x=Opening, y=skips), fun="mean", color="palevioletred1", size=0.8)
```

```{r include=FALSE}
#solder vs skips
solder_plot = ggplot(solder) +
  geom_jitter(aes(x=Solder, y=skips), width=0.1, col="seagreen") +
  theme_bw() +
  labs(title="Thickness of Alloy vs Number of Skips",
       x="Thickness of Alloy",
       y="Number of Skips") +
  stat_summary(aes(x=Solder, y=skips), fun="mean", color="palevioletred1", size=0.8)
```

```{r echo=FALSE, warning=FALSE, fig.width=3.6, fig.height=3}
print(opening_plot)
print(solder_plot)
```

## Part B

```{r echo=FALSE}
#build regression model with skips as outcome and Opening, Solder, and interaction as predictors
solder_model = lm(skips ~ Opening + Solder + Opening:Solder, data=solder) 

solder_table = get_regression_table(solder_model, conf.level=0.95, digits=2) %>%
  select(term, estimate, lower_ci, upper_ci)

kable(solder_table, format = "latex",
      col.names = c("", "Estimate", "Lower CI", "Upper CI"))
```

```{=tex}
\vspace{12pt}
```

## Part C

The baseline number of solder skips on a circuit board that has a small opening and thick alloy is 5.52. The main effect for the medium opening variable is -2.72 skips. This is the effect of a medium opening in isolation. The main effect for the large opening variable is -5.13 skips. This is the effect of a large opening in isolation. The main effect for the thin alloy variable is 11.93 skips. This is the effect of a Thin alloy in isolation. The interaction effect for a medium opening and thin alloy is -10.93. After considering the isolated effects of a medium opening and thin alloy, there is an additional interaction effect to be added.
The interaction effect for a large opening and thin alloy is -9.65. After considering the isolated effects of a large opening and thin alloy, there is an additional interaction effect to be added.


## Part D

The combination that is likely to best minimize the number of skips in the manufacturing process is a large opening size with a thick alloy. Using the regression model, the estimated number of skips with that combination is 0.39, far lower than any other combination.




